Concurrent Priority Queues

### 2014-03-29 presentation/presentation.Rnw
Relaxed slides

\begin{itemize}
\item First SkipList-based \ac{PQ}
\item \fullcite{shavit2000skiplist}
\item Initially lock-based, lock-free variant published in 2008
\item \fullcite{herlihy2012art}
\end{itemize}
\end{frame}

@@

% Intro relaxed

\begin{itemize}
\item
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:relaxed}}
\framesubtitle{\citeauthor{alistarhspraylist}}
% Mention that large PQs are required (weak bound)
\begin{itemize}
\item
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:relaxed}}
\framesubtitle{\citeauthor{wimmer2013data}}
% Essence: Local list, if necessary push to global
\begin{itemize}
\item
\end{itemize}
\end{frame}

>>>>

\begin{itemize}
\item First SkipList-based \ac{PQ}
\item \fullcite{shavit2000skiplist}
\item Initially lock-based (linearizable), lock-free (quiescently consistent)
      variant published in 2008
\item \fullcite{herlihy2012art}
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item Strict \acp{PQ} have inherent bottleneck at minimal element
\item To improve further: $< 1$ \ac{CAS} per \lstinline|DeleteMin|
\item Another approach is to relax semantics, i.e. instead of returning \emph{the} minimal element,
      return one of the $k$ minimal elements
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:relaxed}}
\framesubtitle{\citeauthor{alistarhspraylist}}

\begin{itemize}
\item Relaxed \ac{SL}-based \ac{PQ}, safety properties unclear
\item \lstinline|DeleteMin| returns one of the $O(P \log^3 P)$ elements
\item Degrades to random-remove if the \ac{PQ} is small compared to thread count $P$ (for $P = 80$,
      $P \log^3 P \approx 7000$)
\item \lstinline|DeleteMin| performs random walk: starting at the list head on some level $l$,
      repeatedly follow a randomized number of \lstinline|next[l]| pointers, descend a randomized
      number of levels
\item Parameters are chosen s.t. each element within the walk has approximately equal probability
      of being returned
\item Benchmarks show scalability comparable to a random-remove \lstinline|Delete| up to at
      least 80 threads
\item \fullcite{alistarhspraylist}
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:relaxed}}
\framesubtitle{\citeauthor{wimmer2013data}}
% Essence: Local list, if necessary push to global
\begin{itemize}
\item First relaxed linearizable \acp{PQ}: we discuss the hybrid k-\ac{PQ} which, provides a bound of
      $kP$ missed elements in \lstinline|DeleteMin|
\item Consists of a list of globally visible elements, and per thread: a local item list, and a local
      \ac{PQ}
\item \lstinline|Insert| accesses only the local structures as long as guarantees are not violated,
      otherwise the global list is updated the the local view is synchronized
\item \lstinline|DeleteMin| pops the local queue if it is non-empty; otherwise spy, i.e. copy
      elements from another thread's local list
\item Very good scalability up to 10 threads, further limited gains with rising thread counts
\item \fullcite{wimmer2013data}
\end{itemize}
\end{frame}

\begin{frame}{\nameref{sec:lockfree}}
\framesubtitle{\citeauthor{wimmer2013data}}

\begin{figure}
\includegraphics[width = 0.65\textwidth]{wimmer}
\caption{The \citeauthor{wimmer2013data} hybrid k-\ac{PQ}. (Image source: \cite{wimmer2013data})}
\end{figure}
\end{frame}
